Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-06-16 ms.tex
more comments
comment...
referee 1

\begin{abstract}
  We calculate steady-state, one-dimensional hydrodynamic profiles of
  hot gas in slowly accreting (``quiescent") galactic nuclei for a
  range of central black hole masses $M_{\bullet}$, parameterized gas
  heating rates, and observationally-motivated stellar density
  profiles.  Mass is supplied to the circumnuclear medium by stellar
  winds, while energy is injected primarily by stellar winds,
  supernovae, and black hole feedback.  Analytic estimates are derived
  for the stagnation radius (where the radial velocity of the gas
  passes through zero) and the black hole accretion rate, $\dot{M}$,
  as a function of $M_{\bullet}$ and the gas heating efficiency, the
  latter being related to the star-formation history.  We assess the
  conditions under which radiative instabilities develop in the
  hydrostatic region near the stagnation radius, both in the case of a
  single burst of star formation and for the average star formation
  history predicted by cosmological simulations.  By combining a
  sample of measured nuclear X-ray luminosities, $L_{X}$, of nearby
  quiescent galactic nuclei with our results for
  $\dot{M}(M_{\bullet})$ we address whether the nuclei are consistent
  with accreting in a steady-state, thermally-stable manner for
  radiative efficiencies predicted for radiatively inefficiency
  accretion flows.  We find thermally-stable accretion cannot explain
  the short average growth times of low mass black holes in the local
  universe, which must instead result from gas being fed in from large
  radii, due either to gas inflows or thermal instabilities acting on
  larger, galactic scales.  Our results have implications for attempts
  to constrain the occupation fraction of SMBHs in low mass galaxies
  using the mean $L_{\rm X}-M_{\bullet}$ correlation, as well as the
  predicted diversity of the circumnuclear densities encountered by
  relativistic outflows from tidal disruption events.
\end{abstract}

@@

%The conclusions of
%\citet{Miller+15} are thus sensitive to possible changes in
%the $L_X$-$\Mbh$ relationship at small $\Mbh$.
% AG: In our model we obtain a steeper relationship, but we have difficulty matching the observed relationship above the detection threshold

@@

Another approach to determine the SMBH accretion rate, which we adopt
in this paper, is to directly calculate the density, velocity and
temperature profiles of the CNM using a physically motivated
hydrodynamic model.  Mass is injected into the nuclear environment via
stellar winds, while energy is input from several sources including
stellar winds, supernovae, and AGN feedback
(\citealt{Quataert:2004a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}).
Unlike previous works, which focused primarily on modeling individual
galaxies, here we systematically analyze the CNM properties across a
representative range of galaxy properties, including different SMBH
masses, stellar density profiles, and star formation histories (SFHs).

@@

\begin{table*}
\begin{threeparttable}
\begin{minipage}{18cm}
\caption{Definitions of commonly used variables}
\begin{tabular}{ll}
  \hline
  {Variable} & {Definition} \\
  \hline
  $M_{\bullet}$ & black hole mass \\
  $\tilde{v}_{w}$ & total heating parameter, including minimum heating
  rate from stellar and black hole velocity dispersion \\
  $v_{w}$ & total heating parameter, excluding minimum heating rate from
  velocity dispersion \\
   $\sigma_0$ & stellar velocity dispersion (assumed to be radially
   constant).\\
  $\zeta$ & alternative heating parameter, $\zeta \equiv \sqrt{1 + (v_w/\sigma_0)^2}$ (eq.~[\ref{eq:zetacrit}])\\
  $r_{s}$ & stagnation radius, where gas radial velocity goes to zero \\
  $r_{\rm inf}$ & radius of sphere of influence (eq.~[\ref{eq:rsoi}]) \\
  $r_{\rm b}$ & outer break radius of stellar density profile \\ 
  $r_{\rm Ia}$ & radius interior to which SN Ia are infrequent compared to the dynamical timescale (eq.~[\ref{eq:rIa}]) \\ 
  $\rho_{\star}(r)$ & 3D radial stellar density profile \\
  $\rho(r)$ & gas density of CNM \\
  $M_{\star}(r)$ & total enclosed stellar mass inside radius $r$ \\
  $M_{\rm enc}(r)$ & total enclosed mass inside radius $r$ (SMBH + stars) \\
  $q(r)$ & mass source term due to stellar winds, $q \propto \rho_{\star}$ (eq.~[\ref{eq:q}]) \\
  $\eta$ & parameter setting normalization of mass input from stellar winds (eq.~[\ref{eq:q}]) \\
  $\tau_{\star}$ & age of stellar population, in case of a single burst of star formation \\
  $\Gamma$ & power-law slope of radial stellar surface brightness
  profile interior to the break radius \\
  $\delta$ & power-law slope of the 3D stellar density profile inside of
  the break radius, $\delta \equiv \Gamma+1$.\\
  $\dot{M}$ & accretion rate (not necessarily equal to the SMBH accretion rate)   \\
  $\dot{M}_{\bullet}$ & SMBH accretion rate, $\dot{M}_{\bullet} = \alpha\dot{M}$, where $\alpha < 1$ accounts for outflows from the accretion disk on small scales.   \\
  $\dot{M}_{\rm Ia}$ & maximum mass accretion rate as limited by SN Ia (eq.~[\ref{eq:eddr_Ia}]) \\
  $\dot{M}_{\rm C}$ & equilibrium accretion rate set by Compton heating acting alone (eq.~[\ref{eq:MdotC}]) \\
  $\dot{M}_{\rm TI}$ & maximum accretion rate for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]) \\
  $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ & ratio of external heating (stellar winds, SN Ia, MSPs) to radiative cooling (eq.~[\ref{eq:cooling2}]) \\
  $t_{\rm h}$ & Hubble timescale \\
  $\nu$ & gas density power-law slope at the stagnation radius
  (eq.~[\ref{eq:densSlope}]) \\
  \hline
\label{table:definitions}  
\end{tabular}
%\begin{tablenotes}
%\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

Finally, we note that \citealt{LauerFaber+:2005a} find that $\sim
60\%$ of cusp galaxies and $\sim29\%$ of core galaxies have (generally
unresolved) emission in excess of the inward extrapolation of the
Nuker law.  Indeed, some low mass galaxies with $\Mbh\lsim 10^{8}
\Msun$ possess nuclear star clusters (\citealt{GrahamSpitler:2009a}),
which are not accounted for by our simple parametrization of the
stellar density.  In such cases gas and energy injection could be
dominated by the cluster itself, i.e. concentrated within its own
pc-scale ``break radius" which is much smaller than the outer break in
the older stellar population on much larger scales.  Although our
analysis does not account for such an inner break, we note that for
high heating rates the stagnation radius and concomitant accretion
rate are not sensitive to the break radius.

@@

We first describe analytic estimates of physical quantities, such as
the stagnation radius $\rs$ and the mass accretion rate, the detailed
derivation of which are given in Appendix \ref{app:rs}.

@@

\subsection{Accretion Rate}

The accretion rate towards the SMBH is given by the total mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]), 
\begin{eqnarray}
  \dot{M} &=& 4\pi \int_{0}^{r_{s}}q r^{2}dr = \frac{\eta
    M_{\star}|_{r_{\rm s}}}{\th} = \frac{\eta \Mbh (\rs/\rinf)^{2-\Gamma}}{\th} \nonumber \\
  &\approx&
  \begin{cases}
    4.5 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    3.2 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}            
\end{eqnarray}
where we have assumed $v_{w} \gg \sigma_0$ by adopting equation
(\ref{eq:stag_simple}) for $r_s$.  The resulting Eddington ratio is
given by
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    2.0 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core}, \\
    1.4 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
where $\dot{M}_{\rm edd} = 2.2M_{\bullet,8}M_{\odot}$ yr$^{-1}$ is the
Eddington accretion rate, assuming a radiative efficiency of ten per
cent.  Note the sensitive dependence of the accretion rate on the wind
heating rate.  As will be discussed later,
equation~\eqref{eq:eddr_analytic} is the radial mass inflow rate on
relatively large scales and does not include outflows from the SMBH
accretion disk (e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}),
which may significantly reduce the fraction of $\dot{M}$ that actually
reaches the SMBH.

@@

Even if the hot plasma of the CNM does not condense into cold clouds,
the loss of pressure can temporarily increase the black hole accretion
rate by producing a large-scale cooling flow.  When coupled to
feed-back processes which result from such enhanced accretion, this
can lead to time-dependent limit cycle behavior
(e.g.~\citealt{Ciotti&Ostriker07}; \citealt{Ciotti+10};
\citealt{Yuan&Li11}, \citealt{Gan+14}), which is also inconsistent
with our assumption of a steady, single-phase flow.

@@

Based on our numerical results (\S \ref{sec:numerical}) and the work of \citet{McCourt+12} we define thermally stable flows according to the criterion $\dot{q}_{\rm heat} > 10|\dot{q}_{\rm rad}|$ ($t_{\rm
  cool} > 10t_{\rm ff}$) being satisfied near the stagnation radius.  This condition translates into a
critical minimum heating rate
\begin{align}
v_{w} > v_{\rm TI} \simeq
  \begin{cases}
   280 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   240 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}  \\
%\approx \sigma_0(r_{b}/r_{\rm inf})^{0.5(1-\Gamma)} &, \zeta \ll \zeta_c. 
  \end{cases}
  \label{eq:cooling3}
\end{align}
Equations (\ref{eq:cooling2}) and (\ref{eq:cooling3}) are derived
using expressions for the stagnation radius and gas density in the
high heating limit of $\zeta > \zeta_c = (r_{b}/r_{\rm
  inf})^{0.5(1-\Gamma)}$ (eq.~[\ref{eq:zetacrit}]).  However, for
$\zeta < \zeta_c$, the stagnation radius diverges to the break radius
$r_b$ (Appendix \ref{app:be}).  The quasi-hydrostatic structure that
results in this case greatly increases the gas density, which in
practice renders the flow susceptible to thermal runaway, even if
$v_{w} > v_{\rm TI}$ according to equation (\ref{eq:cooling3}).  In
other words, the true condition for thermal instability  can be
written

@@

The minimum heating rate for thermal stability corresponds to {\it
  maximum} thermally-stable accretion rate. Due to the large scatter
in galaxy structural parameters at a given $\Mbh$, we use
equation~\eqref{eq:cooling3} instead of equation~\eqref{eq:cooling4}
to calculate the minimum heating rate: the $\zeta_c$ criterion should
be evaluated on a galaxy-by-galaxy basis. Thus, the maximum thermally
stable $\Mdot$ (for $\zeta>\zeta_c$) is

@@

\item $^{(a)}$ Break radius of stellar density profile.  $^{(b)}$Accretion
  rate in Eddington units, normalized to a stellar mass input
  parameter $\eta = 0.2$.  $^{(c)}$Ratio of wind heating rate to
  radiative cooling rate at the stagnation radius. $\dagger$Calculated with
  radiative cooling included, assuming mass loss parameter $\eta =
  0.2$.  $\ddagger$Calculated with radiative cooling and conductivity
  included, assuming mass loss parameter $\eta = 0.2$ and conductivity
  saturation parameter $\phi = 0.1$.  Solutions including radiative cooling were performed for cusp galaxies with $v_w=300 $ km s$^{-1}$ and core galaxies with $v_w=600 $ km s$^{-1}$ for $\eta$ = 0.02, 0.2, and 0.6.  Values of $\eta$ resulting in thermally unstable solutions are marked in the final column.  Solutions found to be thermally unstable for all $\eta
  \geq 0.02$ are denoted as {\bf TI}.

@@

Figure~\ref{fig:mdot_mass} shows the SMBH accretion rate for a sample of
our numerical solutions for different values of
$\vwO =$ 300, 600 and 1200 km s$^{-1}$, and for both core
($\Gamma$=0.1) and cusp galaxies ($\Gamma$=0.8).  Shown for comparison
is our simple analytic estimate of $\dot{M}/\dot{M}_{\rm edd}$ from equation
(\ref{eq:eddr_analytic}).  For high wind velocities ($v_{w} \gg
\sigma_0$) the stagnation radius lies well inside the black hole sphere of
influence and our analytic estimate provides a good fit to the
numerical results.  However, for low wind velocities and/or high
$M_{\bullet}$ (large $\sigma_0$), the numerical accretion rate
considerably exceeds the simple analytic estimate as the stagnation radius
diverges to large radii (Fig.~\ref{fig:stag}).

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{mdot_mass.pdf}
  \caption{\label{fig:mdot_mass} Accretion rate $\dot{M}/\dot{M}_{\rm
      edd}$ versus SMBH mass for galaxies in our sample, calculated
    for different values of the wind heating parameter $\vwO =$ 300 km
    s$^{-1}$ ({\it green}), 600 km s$^{-1}$ ({\it orange}), and 1200
    km s$^{-1}$ ({\it blue}).  Squares correspond to cusp galaxies
    ($\Gamma=0.8$), while triangles correspond to cores
    ($\Gamma$=0.1). Green circles correspond to cusp solutions which
    would be thermally unstable.  Thin solid and dot-dashed curves
    correspond to our simple analytic estimates of $\dot{M}/\dot{M}_{\rm
      edd}$ (eq.~[\ref{eq:eddr_analytic}]) for cusp and core galaxies,
    respectively.  Thick curves correspond to the more accurate implicit analytic expression given by equation \eqref{eq:rs2main}.}
\end{figure}

@@

Although most of our solutions neglect thermal conduction and
radiative cooling, these effects are explored explicitly in subset of
simulations.  Dagger symbols in Table \ref{table:models} correspond to
solutions for which we turned on radiative cooling.  For cases which
are far from being thermally unstable when cooling is neglected (e.g.,
$M_{\bullet} = 10^{8}M_{\odot}$, $v_{w} = 600$ km s$^{-1}$, $r_{b} =
100$ pc, $\eta = 0.2$, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| = 76$),
including radiative cooling has little effect on the key properties of
the solution, such as the stagnation radius, mass accretion rate, and
the cooling ratio $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$.  However,
for solutions that are marginally thermally stable, including
radiative losses acts to significantly decrease $\dot{q}_{\rm
  heat}/|\dot{q}_{\rm rad}|$.  We calculated solutions with radiative
cooling for all cusp galaxies with $v_w=300$ km s$^{-1}$ and for all
core galaxies with $v_w=600 $ km s$^{-1}$, each for three different
values of $\eta = 0.02, 0.2, 0.6$, spanning the physical range of
expected mass loss for continuous star formation
(Fig.~\ref{fig:vwSources}).  Solutions found to be unstable for all
values of $\eta$ are marked in boldface as ``TI" in Table
\ref{table:models}, with the unstable values of $\eta$ provided in the
final column.

@@

Key properties of the flow, such as the SMBH accretion rate and the
likelihood of thermal instability depend sensitively on the assumed
heating rate $\propto qv_{w}^{2}$.  This section summarizes individual
heating sources, providing estimates of the value of the effective
wind velocity
\begin{align}
  v_{w} = \sqrt{\frac{2 t_h \dot{e}}{\eta \rhostar}}
  \label{eq:vw_eff}
\end{align}
for each, where $\dot{e}$ is volumetric heating rate.

@@

Feedback from accretion onto the SMBH represents an important source of heating which, however, is also the most difficult to quantify (e.g., \citealt{Brighenti&Mathews03}, \citealt{DiMatteo+05}; \citealt{Kurosawa&Proga09}; \citealt{Fabian12} for a recent review).  A key difference between AGN heating and the other sources discussed thus far is its dependence on the SMBH accretion rate $\dot{M}$, which is itself a function of the heating rate (eq.~[\ref{eq:mdot_analytic}]).

@@

The accretion rate corresponding to a state in which Compton heating
self-regulates the accretion flow is given by substituting equation
(\ref{eq:vC2}) into equation (\ref{eq:eddr_analytic}):
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 2\times 10^{-3} \eta_{0.02}^{0.2}T_{\rm
C,9}^{-0.8}\epsilon_{-2}^{-0.8} M_{\bullet,8}^{0.16}
&, \text{core}\\ 8\times 10^{-4} \eta_{0.02}^{0.3}T_{\rm
C,9}^{-0.7}\epsilon_{-2}^{-0.7} M_{\bullet,8}^{0.14}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotCa}
\end{align}

@@

\subsection{Black Hole Accretion Rates}
\label{sec:mdot}

>>>>

\begin{abstract}
  We calculate steady-state, one-dimensional hydrodynamic profiles of
  hot gas in slowly accreting (``quiescent") galactic nuclei for a
  range of central black hole masses $M_{\bullet}$, parameterized gas
  heating rates, and observationally-motivated stellar density
  profiles.  Mass is supplied to the circumnuclear medium by stellar
  winds, while energy is injected primarily by stellar winds,
  supernovae, and black hole feedback.  Analytic estimates are derived
  for the stagnation radius (where the radial velocity of the gas
  passes through zero) and the large scale gas inflow rate, $\dot{M}$,
  as a function of $M_{\bullet}$ and the gas heating efficiency, the
  latter being related to the star-formation history.  We assess the
  conditions under which radiative instabilities develop in the
  hydrostatic region near the stagnation radius, both in the case of a
  single burst of star formation and for the average star formation
  history predicted by cosmological simulations.  By combining a
  sample of measured nuclear X-ray luminosities, $L_{X}$, of nearby
  quiescent galactic nuclei with our results for
  $\dot{M}(M_{\bullet})$ we address whether the nuclei are consistent
  with accreting in a steady-state, thermally-stable manner for
  radiative efficiencies predicted for radiatively inefficiency
  accretion flows.  We find thermally-stable accretion cannot explain
  the short average growth times of low mass black holes in the local
  universe, which must instead result from gas being fed in from large
  radii, due either to gas inflows or thermal instabilities acting on
  larger, galactic scales.  Our results have implications for attempts
  to constrain the occupation fraction of SMBHs in low mass galaxies
  using the mean $L_{\rm X}-M_{\bullet}$ correlation, as well as the
  predicted diversity of the circumnuclear densities encountered by
  relativistic outflows from tidal disruption events.
\end{abstract}

@@

@@

Another approach to determine the SMBH accretion rate, which we adopt
in this paper, is to directly calculate the density, velocity and
temperature profiles of the CNM using a physically motivated
hydrodynamic model.  Mass is injected into the nuclear environment via
stellar winds, while energy is input from several sources including
stellar winds, supernovae, and AGN feedback
(\citealt{Quataert:2004a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}).
Unlike previous works, which focused primarily on modeling individual
galaxies, here we model the CNM properties across a
representative range of galaxy properties, including different SMBH
masses, stellar density profiles, and star formation histories (SFHs).

@@

\begin{table*}
\begin{threeparttable}
\begin{minipage}{18cm}
\caption{Definitions of commonly used variables}
\begin{tabular}{ll}
  \hline
  {Variable} & {Definition} \\
  \hline
  $M_{\bullet}$ & black hole mass \\
  $\tilde{v}_{w}$ & total heating parameter, including minimum heating
  rate from stellar and black hole velocity dispersion \\
  $v_{w}$ & total heating parameter, excluding minimum heating rate from
  velocity dispersion \\
   $\sigma_0$ & stellar velocity dispersion (assumed to be radially
   constant).\\
  $\zeta$ & alternative heating parameter, $\zeta \equiv \sqrt{1 + (v_w/\sigma_0)^2}$ (eq.~[\ref{eq:zetacrit}])\\
  $r_{s}$ & stagnation radius, where gas radial velocity goes to zero \\
  $r_{\rm inf}$ & radius of sphere of influence (eq.~[\ref{eq:rsoi}]) \\
  $r_{\rm b}$ & outer break radius of stellar density profile \\ 
  $r_{\rm Ia}$ & radius interior to which SN Ia are infrequent compared to the dynamical timescale (eq.~[\ref{eq:rIa}]) \\ 
  $\rho_{\star}(r)$ & 3D radial stellar density profile \\
  $\rho(r)$ & gas density of CNM \\
  $M_{\star}(r)$ & total enclosed stellar mass inside radius $r$ \\
  $M_{\rm enc}(r)$ & total enclosed mass inside radius $r$ (SMBH + stars) \\
  $q(r)$ & mass source term due to stellar winds, $q \propto \rho_{\star}$ (eq.~[\ref{eq:q}]) \\
  $\eta$ & parameter setting normalization of mass input from stellar winds (eq.~[\ref{eq:q}]) \\
  $\tau_{\star}$ & age of stellar population, in case of a single burst of star formation \\
  $\Gamma$ & power-law slope of radial stellar surface brightness
  profile interior to the break radius \\
  $\delta$ & power-law slope of the 3D stellar density profile inside of
  the break radius, $\delta \equiv \Gamma+1$.\\
  $\dot{M}$ & large scale inflow rate (not necessarily equal to the SMBH accretion rate)   \\
  $\dot{M}_{\bullet}$ & SMBH accretion rate, $\dot{M}_{\bullet} = \alpha\dot{M}$, where $\alpha < 1$ accounts for outflows from the accretion disk on small scales.   \\
  $\dot{M}_{\rm Ia}$ & maximum accretion rate as limited by SN Ia (eq.~[\ref{eq:eddr_Ia}]) \\
  $\dot{M}_{\rm C}$ & equilibrium inflow rate set by Compton heating acting alone (eq.~[\ref{eq:MdotC}]) \\
  $\dot{M}_{\rm TI}$ & maximum accretion rate for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]) \\
  $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ & ratio of external heating (stellar winds, SN Ia, MSPs) to radiative cooling (eq.~[\ref{eq:cooling2}]) \\
  $t_{\rm h}$ & Hubble timescale \\
  $\nu$ & gas density power-law slope at the stagnation radius
  (eq.~[\ref{eq:densSlope}]) \\
  \hline
\label{table:definitions}  
\end{tabular}
%\begin{tablenotes}
%\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

Finally, we note that \citealt{LauerFaber+:2005a} find that $\sim
60\%$ of cusp galaxies and $\sim29\%$ of core galaxies have (generally
unresolved) emission in excess of the inward extrapolation of the
Nuker law.  Indeed, some low mass galaxies with $\Mbh\lsim 10^{8}
\Msun$ possess nuclear star clusters (\citealt{GrahamSpitler:2009a}),
which are not accounted for by our simple parametrization of the
stellar density.  In such cases gas and energy injection could be
dominated by the cluster itself, i.e. concentrated within its own
pc-scale ``break radius" which is much smaller than the outer break in
the older stellar population on much larger scales.  Although our
analysis does not account for such an inner break, we note that for
high heating rates the stagnation radius and concomitant inflow
rate are not sensitive to the break radius.

@@

We first describe analytic estimates of physical quantities, such as
the stagnation radius $\rs$ and the mass inflow rate, the detailed
derivation of which are given in Appendix \ref{app:rs}.

@@

\subsection{Inflow Rate}

The large scale inflow rate towards the SMBH is given by the total
mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]),
\begin{eqnarray}
  \dot{M} &=& 4\pi \int_{0}^{r_{s}}q r^{2}dr = \frac{\eta
    M_{\star}|_{r_{\rm s}}}{\th} = \frac{\eta \Mbh (\rs/\rinf)^{2-\Gamma}}{\th} \nonumber \\
  &\approx&
  \begin{cases}
    4.5 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    3.2 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}            
\end{eqnarray}
where we have assumed $v_{w} \gg \sigma_0$ by adopting equation
(\ref{eq:stag_simple}) for $r_s$.  The resulting Eddington ratio is
given by
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    2.0 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core}, \\
    1.4 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
where $\dot{M}_{\rm edd} = 2.2M_{\bullet,8}M_{\odot}$ yr$^{-1}$ is the
Eddington accretion rate, assuming a radiative efficiency of ten per
cent.  Note the sensitive dependence of the inflow rate on the wind
heating rate.  As will be discussed later,
equation~\eqref{eq:eddr_analytic} is the radial mass inflow rate on
relatively large scales and does not include outflows from the SMBH
accretion disk (e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}),
which may significantly reduce the fraction of $\dot{M}$ that actually
reaches the SMBH. Thus we distinguish between the large scale inflow
rate, $\dot{M}$, and the accretion rate onto the black hole,
$\dot{M}_{\bullet}$.

@@

Even if the hot plasma of the CNM does not condense into cold clouds,
the loss of pressure can temporarily increase the inflow and accretion
rates by producing a large-scale cooling flow.  When coupled to
feed-back processes which result from such enhanced accretion, this
can lead to time-dependent limit cycle behavior
(e.g.~\citealt{Ciotti&Ostriker07}; \citealt{Ciotti+10};
\citealt{Yuan&Li11}, \citealt{Gan+14}), which is also inconsistent
with our assumption of a steady, single-phase flow.

@@

Based on our numerical results (\S \ref{sec:numerical}) and the work of \citet{McCourt+12} we define thermally stable flows according to the criterion $\dot{q}_{\rm heat} > 10|\dot{q}_{\rm rad}|$ ($t_{\rm
  cool} > 10t_{\rm ff}$) being satisfied near the stagnation radius.  This condition translates into a
critical minimum heating rate
\begin{align}
v_{w} > v_{\rm TI} \simeq
  \begin{cases}
   280 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   240 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}  \\
  \end{cases}
  \label{eq:cooling3}
\end{align}
Equations (\ref{eq:cooling2}) and (\ref{eq:cooling3}) are derived
using expressions for the stagnation radius and gas density in the
high heating limit of $\zeta > \zeta_c = (r_{b}/r_{\rm
  inf})^{0.5(1-\Gamma)}$ (eq.~[\ref{eq:zetacrit}]).  However, for
$\zeta < \zeta_c$, the stagnation radius diverges to the break radius
$r_b$ (Appendix \ref{app:be}).  The quasi-hydrostatic structure that
results in this case greatly increases the gas density, which in
practice renders the flow susceptible to thermal runaway, even if
$v_{w} > v_{\rm TI}$ according to equation (\ref{eq:cooling3}).  In
other words, the true condition for thermal instability  can be
written

@@

The minimum heating rate for thermal stability corresponds to {\it
  maximum} thermally-stable inflow rate. Due to the large scatter
in galaxy structural parameters at a given $\Mbh$, we use
equation~\eqref{eq:cooling3} instead of equation~\eqref{eq:cooling4}
to calculate the minimum heating rate: the $\zeta_c$ criterion should
be evaluated on a galaxy-by-galaxy basis. Thus, the maximum thermally
stable $\Mdot$ (for $\zeta>\zeta_c$) is

@@

\item $^{(a)}$ Break radius of stellar density profile.
  $^{(b)}$Inflow rate in Eddington units, normalized to a stellar mass
  input parameter $\eta = 0.2$.  $^{(c)}$Ratio of wind heating rate to
  radiative cooling rate at the stagnation radius. $\dagger$Calculated
  with radiative cooling included, assuming mass loss parameter $\eta
  = 0.2$.  $\ddagger$Calculated with radiative cooling and
  conductivity included, assuming mass loss parameter $\eta = 0.2$ and
  conductivity saturation parameter $\phi = 0.1$.  Solutions including
  radiative cooling were performed for cusp galaxies with $v_w=300 $
  km s$^{-1}$ and core galaxies with $v_w=600 $ km s$^{-1}$ for $\eta$
  = 0.02, 0.2, and 0.6.  Values of $\eta$ resulting in thermally
  unstable solutions are marked in the final column.  Solutions found
  to be thermally unstable for all $\eta \geq 0.02$ are denoted as
  {\bf TI}.

@@

Figure~\ref{fig:mdot_mass} shows the inflow rate for a sample of
our numerical solutions for different values of
$\vwO =$ 300, 600 and 1200 km s$^{-1}$, and for both core
($\Gamma$=0.1) and cusp galaxies ($\Gamma$=0.8).  Shown for comparison
is our simple analytic estimate of $\dot{M}/\dot{M}_{\rm edd}$ from equation
(\ref{eq:eddr_analytic}).  For high wind velocities ($v_{w} \gg
\sigma_0$) the stagnation radius lies well inside the black hole sphere of
influence and our analytic estimate provides a good fit to the
numerical results.  However, for low wind velocities and/or high
$M_{\bullet}$ (large $\sigma_0$), the numerical accretion rate
considerably exceeds the simple analytic estimate as the stagnation radius
diverges to large radii (Fig.~\ref{fig:stag}).

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{mdot_mass.pdf}
  \caption{\label{fig:mdot_mass} Inflow rate $\dot{M}/\dot{M}_{\rm
      edd}$ versus SMBH mass for galaxies in our sample, calculated
    for different values of the wind heating parameter $\vwO =$ 300 km
    s$^{-1}$ ({\it green}), 600 km s$^{-1}$ ({\it orange}), and 1200
    km s$^{-1}$ ({\it blue}).  Squares correspond to cusp galaxies
    ($\Gamma=0.8$), while triangles correspond to cores
    ($\Gamma$=0.1). Green circles correspond to cusp solutions which
    would be thermally unstable.  Thin solid and dot-dashed curves
    correspond to our simple analytic estimates of $\dot{M}/\dot{M}_{\rm
      edd}$ (eq.~[\ref{eq:eddr_analytic}]) for cusp and core galaxies,
    respectively.  Thick curves correspond to the more accurate implicit analytic expression given by equation \eqref{eq:rs2main}.}
\end{figure}

@@

Although most of our solutions neglect thermal conduction and
radiative cooling, these effects are explored explicitly in subset of
simulations.  Dagger symbols in Table \ref{table:models} correspond to
solutions for which we turned on radiative cooling.  For cases which
are far from being thermally unstable when cooling is neglected (e.g.,
$M_{\bullet} = 10^{8}M_{\odot}$, $v_{w} = 600$ km s$^{-1}$, $r_{b} =
100$ pc, $\eta = 0.2$, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| = 76$),
including radiative cooling has little effect on the key properties of
the solution, such as the stagnation radius, mass inflow rate, and
the cooling ratio $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$.  However,
for solutions that are marginally thermally stable, including
radiative losses acts to significantly decrease $\dot{q}_{\rm
  heat}/|\dot{q}_{\rm rad}|$.  We calculated solutions with radiative
cooling for all cusp galaxies with $v_w=300$ km s$^{-1}$ and for all
core galaxies with $v_w=600 $ km s$^{-1}$, each for three different
values of $\eta = 0.02, 0.2, 0.6$, spanning the physical range of
expected mass loss for continuous star formation
(Fig.~\ref{fig:vwSources}).  Solutions found to be unstable for all
values of $\eta$ are marked in boldface as ``TI" in Table
\ref{table:models}, with the unstable values of $\eta$ provided in the
final column.

@@

Key properties of the flow, such as the mass inflow rate and the
likelihood of thermal instability depend sensitively on the assumed
heating rate $\propto qv_{w}^{2}$.  This section summarizes individual
heating sources, providing estimates of the value of the effective
wind velocity
\begin{align}
  v_{w} = \sqrt{\frac{2 t_h \dot{e}}{\eta \rhostar}}
  \label{eq:vw_eff}
\end{align}
for each, where $\dot{e}$ is volumetric heating rate.

@@

Feedback from accretion onto the SMBH represents an important source
of heating which, however, is also the most difficult to quantify
(e.g., \citealt{Brighenti&Mathews03}, \citealt{DiMatteo+05};
\citealt{Kurosawa&Proga09}; \citealt{Fabian12} for a recent review).
A key difference between AGN heating and the other sources discussed
thus far is its dependence on the SMBH accretion rate
$\dot{M}_{\bullet}$, which is itself a function of the heating rate
(eq.~[\ref{eq:mdot_analytic}]).

@@

The inflow rate corresponding to a state in which Compton heating
self-regulates the accretion flow is given by substituting equation
(\ref{eq:vC2}) into equation (\ref{eq:eddr_analytic}):
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 2\times 10^{-3} \eta_{0.02}^{0.2}T_{\rm
C,9}^{-0.8}\epsilon_{-2}^{-0.8} M_{\bullet,8}^{0.16}
&, \text{core}\\ 8\times 10^{-4} \eta_{0.02}^{0.3}T_{\rm
C,9}^{-0.7}\epsilon_{-2}^{-0.7} M_{\bullet,8}^{0.14}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotCa}
\end{align}

@@

\subsection{Inflow and Black Hole Accretion Rates}
\label{sec:mdot}
